---
title: Language distinctiveness
subtitle: difference
author: Molly Lewis 
date: "`r Sys.Date()`"
output: 
  html_document:
    toc_float: no
    number_sections: no
    code_folding: hide
    toc: no
---

```{r setup, include = F}
# load packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(tidyboot)
library(cowplot)
library(here)
library(broom)
library(glue)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, tidy = F, cache = F)
```

## scores
```{r}
SCORE_METADATA <-  here("data/raw/models/all_model/merged_metadata.csv")
meta_data <- read_csv(SCORE_METADATA)

score_means_text <- meta_data %>%
  summarize(mean_score = mean(score), 
            sd_score = sd(score)) %>%
  mutate_if(is.numeric, round, 2)

kable(score_means_text)

glue("$M$ = {score_means_text$mean_score}, $SD$ = {score_means_text$sd_score}") 
```


## Disinctiveness by score group{.tabset}

### All
```{r}
LANG_NAME_PATH <- here("data/processed/lang_names/ets_to_google_langcodes_complete.csv")

langs_clean <- read_csv(LANG_NAME_PATH)


DISTANCE_INFILE_ALL <- here("analyses/01_distinctiveness/data/sampled_essay_distances_all.csv")

mean_language_distances_all <- read_csv(DISTANCE_INFILE_ALL, 
                                        col_names = c("language", "score_group", "distance_type", "distance", "n", "sample_id")) %>%
  mutate_if(is.character, as.factor)

dists_wide_all <- mean_language_distances_all %>%
  select(-n) %>%
  spread(distance_type, distance) %>%
  mutate(same_diff_distance = same_language - diff_language) %>%
  select(-diff_language, -same_language)

lang_means_all <- dists_wide_all %>%
  group_by(language) %>%
  tidyboot_mean(column = same_diff_distance, nboot = 1000)

lang_means_all %>%
  left_join(langs_clean %>% 
            mutate(ETS_lang_name = toupper(ETS_lang_name)), 
          by = c("language" = "ETS_lang_name")) %>%
  ggplot(aes(x = lang_name3, y = mean))   +
  geom_bar(stat = "identity", position = "dodge") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),
                 position = position_dodge(width = 1)) +
  scale_y_continuous(name = "Semantic distinctiveness\n(within - between semantic distance)") +
  scale_fill_manual(values = c( "#fb9a99", "#a6cee3"), name = "Essay Score") + 
  xlab("Language") +
  theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        legend.position = c(.1, 0.8),
        axis.line = element_line(size = 1.2),
        axis.ticks = element_line(size = 1),
        legend.background = element_blank())
```


#### Stats
Descriptives
```{r}
# take mean across runs
language_means <- dists_wide_all %>%
  group_by(language) %>%
  summarize(same_diff_distance = mean(same_diff_distance)) 

# take mean across langauges
lang_means_all_text <- language_means %>%
  summarize(mean = mean(same_diff_distance),
            sd = sd(same_diff_distance)) %>%
  mutate_all(round, 3) 

kable(lang_means_all_text)
```

Is each language > 0?
```{r}
MU <- 0
t_test_all_text  <- t.test(language_means %>% 
         pull(same_diff_distance),
       mu = MU) %>%
  tidy() %>%
  mutate_at(vars(estimate, statistic), round, 2) 

kable(t_test_all_text)

glue("$M$ = {lang_means_all_text$mean}, $SD$ = {lang_means_all_text$sd}; $t$({t_test_all_text$parameter}) = {t_test_all_text$statistic}, $p$ $<$ .00001")
```

Effect size - to do
```{r, eval = F}
psych::cohen.d(language_means %>% data.frame(), "same_diff_distance")
cohen.d(d = language_means$same_diff_distance,  f = NA, formula = "y ~ .", mu = 0) 
  mutate(id = 1:n()) %>%
  select(id, local, global) %>%
  pivot_longer(cols = 2:3) %>%
  group_by(name) %>%
  summarize(m = mean(value),
         sd = sd(value),
         n = n())
  
wiki_es <- mes(es_data_wiki %>% filter(name == "local") %>% pull(m),
    es_data_wiki %>% filter(name == "global") %>% pull(m),
    es_data_wiki %>% filter(name == "local") %>% pull(sd),
    es_data_wiki %>% filter(name == "global") %>% pull(sd),
    es_data_wiki %>% filter(name == "local") %>% pull(n),
    es_data_wiki %>% filter(name == "global") %>% pull(n),
    verbose = F)
glue("Wiki: $d$ = {wiki_es$d} [{wiki_es$l.d}, {wiki_es$u.d}]")
```


#####  Testing for normality: Kolmogorov-Smirnov test.

Null hypothesis: test distribution is normal
```{r}
ks.test(language_means$same_diff_distance, "pnorm", 
        mean=mean(language_means$same_diff_distance), 
        sd=sd(language_means$same_diff_distance))
```

##### Non-parameteric t-test: Wilcoxon signed-rank test
null hypothesis: two means the same

```{r}
wilcox_all <- wilcox.test(language_means$same_diff_distance, mu = MU)
wilcox_all
glue("$W$ = {wilcox_all$statistic}, $p$ $<$ .00001")
```


### Low vs.high
```{r paths}
DISTANCE_INFILE <- here("analyses/01_distinctiveness/data/sampled_essay_distances_high_low.csv")
FIG_PATH <- here("analyses/01_distinctiveness/results/distinctiveness_diff_fig.pdf")
```

Main plot
```{r, fig.width = 10}
mean_language_distances <- read_csv(DISTANCE_INFILE,
                                         col_names = c("language", "score_group", "distance_type", "distance", "n", "sample_id"))  

dists_wide <- mean_language_distances %>%
  select(-n) %>%
  group_by(score_group) %>%
  spread(distance_type, distance) %>%
  mutate(same_diff_distance = same_language - diff_language) %>%
  select(-diff_language, -same_language)

lang_means <- dists_wide %>%
  group_by(language, score_group) %>%
  tidyboot_mean(column = same_diff_distance, nboot = 1000)

main_plot <- lang_means %>%
  left_join(langs_clean %>% 
            mutate(ETS_lang_name = toupper(ETS_lang_name)), 
          by = c("language" = "ETS_lang_name")) %>%
  ggplot(aes(x = lang_name3, y = mean, group = score_group, fill = score_group))   +
  geom_bar(stat = "identity", position = "dodge") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),
                 position = position_dodge(width = 1)) +
  scale_y_continuous(name = "Semantic distinctiveness\n(within - between semantic distance)",
                     limits = c(0,.05)) +
  scale_fill_manual(values = c( "#fb9a99", "#a6cee3"), name = "Essay Score") + 
  xlab("Language") +
  theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        legend.position = c(.1, 0.8),
        axis.line = element_line(size = 1.2),
        axis.ticks = element_line(size = 1),
        legend.background = element_blank())
```

Histogram inset
```{r}
inset_plot <- ggplot(meta_data, aes(x = score)) +
  geom_histogram(binwidth = 1) +
  xlab("Essay Score") +
  ylab("N Essays") +
  theme_classic(base_size = 10) 
```

```{r}
plot_with_inset <-
  ggdraw() +
  draw_plot(main_plot) +
  draw_plot(inset_plot, x = 0.83, y = .75, width = .14, height = .2)

plot_with_inset
```

```{r paper_fig, eval = F}
pdf(FIG_PATH, width = 14)
plot_with_inset
dev.off()
```

#### Stats

##### Descriptives
```{r}
language_means_hl <- dists_wide %>%
  group_by(language, score_group) %>%
  summarize(same_diff_distance = mean(same_diff_distance)) %>%
  mutate(score_group = fct_rev(score_group))

language_means_hl_text <- language_means_hl %>%
  group_by(score_group) %>%
  summarize(mean = mean(same_diff_distance),
            sd = sd(same_diff_distance)) %>%
  mutate_if(is.numeric, round, 3) 

kable(language_means_hl_text)
```

##### paired t-tests: For each language pair is low > high?
```{r}
t_test_hl_text <- t.test(same_diff_distance ~ score_group, 
       data = language_means_hl, paired = T) %>%
  tidy() %>%
  mutate_at(vars(estimate, statistic), round, 2) 

kable(t_test_hl_text)

glue("$M$ = {language_means_hl_text %>% filter(score_group == 'low') %>% pull(mean)}, $SD$ = {language_means_hl_text %>% filter(score_group == 'low') %>% pull(sd)}") 

glue("$M$ = {language_means_hl_text %>% filter(score_group == 'high') %>% pull(mean)}, $SD$ = {language_means_hl_text %>% filter(score_group == 'high') %>% pull(sd)}; $t$({t_test_hl_text$parameter}) = {t_test_hl_text$statistic}, $p$ $<$ .001")

```

#####  Testing for normality: Kolmogorov-Smirnov test.

Null hypothesis: test distribution is normal
```{r}
low_data <- language_means_hl %>%
  filter(score_group == "low") %>%
  pull(same_diff_distance)

high_data <- language_means_hl %>%
  filter(score_group == "high") %>%
  pull(same_diff_distance)
  
ks.test(low_data, "pnorm", mean=mean(low_data), sd=sd(low_data))
ks.test(high_data, "pnorm", mean=mean(high_data), sd=sd(high_data))
```

##### Non-parameteric t-test: Wilcoxin signed-rank test
null hypothesis: two samples come from same distribution

```{r}
wilcox_paired <- wilcox.test(low_data, high_data, paired = TRUE)
wilcox_paired
glue("$W$ = {wilcox_paired$statistic}, $p$ $<$ .001")
```
